\textbf{Appendix}\\ A Three-Step Approach for Assessing Landscape Connectivity via Simulated Dispersal: African Wild Dog Case Study

### 2022-11-18 Chapter_1/04_Manuscript/Manuscript_Support.tex
Revisions done

@@

%	Appendix S4: All Movement Models

@@

%------------------------------------------------------------------------------
%	Appendix S5: Most Parsimonious Movement Models
%------------------------------------------------------------------------------
\newpage
\section{Movement Model}
\begin{table}[htbp]
 \begin{center}
  \caption{Most parsimonious movement model for dispersing wild dogs. The model
   consists of a movement kernel, a habitat kernel, and their interactions. The
   movement kernel describes preferences with regards to movement behavior,
   whereas the habitat kernel describes preferences with respect to habitat
   conditions. Interactions between the two kernels indicate that movement
   preferences are contingent on habitat conditions. Note that all covariates
   were standardized to a mean of zero and standard deviation of 1. Plots to aid
   with the interpretation of this model are given in Appendix S2.}
  \label{MovementModelNumbers}
  \resizebox{\textwidth}{!} {
   \begin{threeparttable}
    \begin{tabular}{llrrrc}
     \toprule
     Kernel & Covariate                                       & Coefficient & SE    & p-value     & Sign. \\
     \midrule
     \multirow{5}{*}{Habitat Kernel}
            & Water                                           & -0.546      & 0.112 & \(<\) 0.001 & ***   \\
            & DistanceToWater \textsuperscript{0.5}           & -0.390      & 0.231 & 0.092       & *     \\
            & Woodland                                        & -0.364      & 0.086 & \(<\) 0.001 & ***   \\
            & Shrubs/Grassland                                & 0.288       & 0.092 & 0.002       & ***   \\
            & HumanInfluence                                  & -0.535      & 0.229 & 0.019       & **    \\
     \hdashline
     \multirow{6}{*}{Movement Kernel}
            & sl                                              & 0.075       & 0.037 & 0.042       & **    \\
            & cos(ta)                                         & 0.105       & 0.031 & 0.001       & ***   \\
            & log(sl)                                         & 0.146       & 0.051 & 0.004       & ***   \\
            & cos(ta) : sl                                    & 0.049       & 0.026 & 0.064       & *     \\
            & cos(ta) : log(sl)                               & 0.076       & 0.026 & 0.003       & ***   \\
            & sl : LowActivity                                & -0.917      & 0.113 & \(<\) 0.001 & ***   \\
     \hdashline
     \multirow{5}{*}{Interactions}
            & sl : Water                                      & -0.305      & 0.076 & \(<\) 0.001 & ***   \\
            & sl : Woodland                                   & -0.089      & 0.039 & 0.023       & **    \\
            & sl : Shrubs/Grassland                           & 0.124       & 0.058 & 0.032       & **    \\
            & sl : DistanceToWater \textsuperscript{0.5}      & -0.058      & 0.031 & 0.056       & *     \\
            & cos(ta) : HumanInfluence                        & -0.040      & 0.022 & 0.070       & *     \\
            & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063       & 0.026 & 0.017       & **    \\
     \bottomrule
    \end{tabular}
    \begin{tablenotes}
     \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
      \quad *** \(p < 0.01\)}
    \end{tablenotes}
   \end{threeparttable}
  }
 \end{center}
\end{table}

%------------------------------------------------------------------------------
%	Appendix S6: Movement Model interpretation
%------------------------------------------------------------------------------
\newpage
% \newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
\section{Movement Model Interpretation}
To ease with the interpretation of the most parsimonious movement model, we
followed recommendations published in \cite{Fieberg.2021} and produced a series
of plots highlighting how the habitat and movement kernel depended on covariate
values (\Cref{Interpretation}). To visualize the movement kernel and its
interactions with other covariates, we used model estimates and updated our
tentative distribution parameters for turning angles (von Mises distribution
with concentration \(\kappa = 0\)) and step lengths (gamma distribution with
scale \(\theta\) = 6'308 and shape \(k\) = 0.37) by applying the function
\textsf{update\_vonmises()} from the R-package \textsf{amt} \citep{Amt.2019}.
This allowed us to compute probability densities of turning angles and step
lengths under varying values of the associated covariates, while holding all
other covariates constant (\Cref{Interpretation}, a1-a8). Moreover, we
investigated the habitat kernel by computing relative selection strengths (RSS)
between a set of steps where values of the covariate of interest was varied to a
reference step where the covariate value was fixed to its centered value. To
illustrate model uncertainty, we also generated large-sample confidence
intervals using standard errors associated with each model estimate
(\Cref{Interpretation}, b1-b5).

@@

%------------------------------------------------------------------------------
%	Appendix S7: Convergence
%------------------------------------------------------------------------------
\newpage
\section{Convergence}
\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{99_Convergence}
    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5
    km) distributed randomly across the study area. The relative traversal
    frequency is plotted against the number of simulated individuals to
    visualize how quickly the metric converges to a steady state. (a) Replicated
    (100 times) relative traversal frequencies across three randomly chosen
    checkpoints as well as the corresponding 95\% prediction interval (PI). (b)
    Averaged relative traversal frequency across all checkpoints and replicates
    including a 95\% PI. (c) Width of the PI in relation to the number of
    simulated dispersers.}
    \label{Convergence}
  \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S8: Heatmaps
%------------------------------------------------------------------------------
\newpage
\newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
\section{Heatmaps in Relation to the Number of Simulated Steps}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = 0.86\textwidth]{99_HeatmapsIndividual.png}
  \caption{Heatmaps produced when considering 125, 500, and 2000 simulated
   steps, respectively. The left panel (a1, a2, a3) was generated based on
   simulations initiated within the main study area, the right panel (b1, b2,
   b3) was generated based on simulations initiated within the buffer area. To
   produce the heatmap presented in the main manuscript (Figure 5), we tallied
   the values from maps a3 and b3.}
  \label{Heatman}
 \end{center}
\end{figure}
\restoregeometry

%------------------------------------------------------------------------------
%	Appendix S9: Evolution of Betweenness Maps
%------------------------------------------------------------------------------
\newpage
\newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
\section{Betweenness Maps in Relation to the Number of Simulated Steps}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = 0.86\textwidth]{99_BetweennessIndividual.png}
  \caption{Maps of betweenness scores produced when considering (a) 125, (b)
   500, (c) and 2000 simulated steps, respectively. A high betweenness score
   indicates that the respective area has a high importance for linking other
   regions in the study area.}
  \label{Betweenness}
 \end{center}
\end{figure}
\restoregeometry

%------------------------------------------------------------------------------
%	Appendix S11: Map Comparison
%------------------------------------------------------------------------------
\newpage
\section{Comparison of Traversal Frequencies and Betweenness Scores Inside and Outside KAZA-TFCA}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = \textwidth]{99_MapComparison.png}
  \caption{Comparison of values from the heatmap and betweenness map inside
  (blue) and outside (orange) the KAZA-TFCA borders.}
  \label{MapComparison}
 \end{center}
\end{figure}

@@

%------------------------------------------------------------------------------
%	Appendix S12: Areas Reached
%------------------------------------------------------------------------------
\newpage
\section{Dispersal into other National Parks}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = \textwidth]{99_AreasReached.png}
  \caption{Relative number of simulated dispersal trajectories that successfully
  moved from one national park into another that is at least as far away as
  indicated on the x-axis. Percentages are given in relation to the number of
  simulated individuals from the national parks in the respective countries. For
  example, over 85\% of all individuals originating from a national park in
  Angola moved from their natal national park into another one. However, the
  percentage gradually decreases as only national parks at higher euclidean
  distances are considered.}
  \label{AreasReached}
 \end{center}
\end{figure}

>>>>

%------------------------------------------------------------------------------
%	Appendix S4: Betweenness Approaches
%------------------------------------------------------------------------------
\newpage
\section{Betweenness Maps}
To generate betweenness maps, we first overlaid the study area with a regular
grid with a resolution of 2.5 km x 2.5 km per grid cell. The grid cells were
labelled with unique numbers, which allowed us to then generate a ``visitation
history'' for each simulated trajectory, showing through which grid cells each
simulated invidual moved. Based on this history, we tallied the number of
transitions that occured from one cell to another. Ultimately, the so generated
list of transitions served to generate a network and to compute betweenness
scores. The resolution of the overlaid grid thus dictates the resolution at
which betweenness can be mapped and has important implications for the way in
which cell-transitions need to be determined. In the simplest case,
cell-transitions can be determined by only looking at the start and endpoint of
each step. If it ends in a different grid-cell than it started, a transition has
occured. This approach is computationally simple as it only requires a point
extraction (at the start and endpoint of each step). However, relying on a point
extraction has the disadvantage that transitions between non-adjacent cells can
occur if a single step stretches across multiple grid-cells. Especially when the
resolution of the grid is high, this problem becomes more pertinent. As an
alternative, one could not only consider the start and endpoint of each step,
but the entire line along the step, i.e. conduct a line extraction. However,
this requires that extracted values (i.e. grid-cell labels) are ordered in
accordance to the direction of movement. In R, such an extraction can be
achieved e.g. using \textsf{raster::extract(raster, line, along = T)}, which is
computationally very demanding and therefore often an unviable solution. Here,
we employed yet another approach and generated interpolated coordinates along
each simulated step. We then determined cell transition using those interpolated
coordinates. Since this approach still uses a point extraction, it is
considerably faster while still enabling to detect cell-transitions at greater
resolution. In \Cref{BetweennessApproaches}, we use simulated movement paths to
compare betweenness maps that are based on a (1) point extraction, a (2) line
extraction, and a (3) point extraction with interpolated coordinates. As you can
see, the approach using interpolated coordinates approximates the results from
the line extraction very closely, even at higher resolutions. A detailed
discussion about this is can also be found in \citep{BastilleRousseau.2018}.

\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width = \textwidth]{99_BetweennessApproaches.png}
  \caption{We simulated five trajectories where individuals moved through a
  narrow corridor. We then overlaid those simulations with a coarse (a1) and
  detailed (b1) grid that was used to compute betweenness scores. This required
  us to determine cell-transitions, which we did using three competing
  approaches. Once using a point extraction (a2 and b2), once using a line
  extraction (a3 and b3), and once using a point extraction with interpolated
  coordinates (a4 and b4). In brackets are the average computation times $\pm$
  SD as computed from 10 repetitions.}
  \label{BetweennessApproaches}
 \end{center}
\end{figure}

@@

%	Appendix S5: All Movement Models

@@

%------------------------------------------------------------------------------
%	Appendix S6: Most Parsimonious Movement Models
%------------------------------------------------------------------------------
\newpage
\section{Movement Model}
\begin{table}[htbp]
 \begin{center}
  \caption{Most parsimonious movement model for dispersing wild dogs. The model
   consists of a movement kernel, a habitat kernel, and their interactions. The
   movement kernel describes preferences with regards to movement behavior,
   whereas the habitat kernel describes preferences with respect to habitat
   conditions. Interactions between the two kernels indicate that movement
   preferences are contingent on habitat conditions. Note that all covariates
   were standardized to a mean of zero and standard deviation of 1. Plots to aid
   with the interpretation of this model are given in Appendix S2.}
  \label{MovementModelNumbers}
  \resizebox{\textwidth}{!} {
   \begin{threeparttable}
    \begin{tabular}{llrrrc}
     \toprule
     Kernel & Covariate                                       & Coefficient & SE    & p-value     & Sign. \\
     \midrule
     \multirow{5}{*}{Habitat Kernel}
            & Water                                           & -0.546      & 0.112 & \(<\) 0.001 & ***   \\
            & DistanceToWater \textsuperscript{0.5}           & -0.390      & 0.231 & 0.092       & *     \\
            & Woodland                                        & -0.364      & 0.086 & \(<\) 0.001 & ***   \\
            & Shrubs/Grassland                                & 0.288       & 0.092 & 0.002       & ***   \\
            & HumanInfluence                                  & -0.535      & 0.229 & 0.019       & **    \\
     \hdashline
     \multirow{6}{*}{Movement Kernel}
            & sl                                              & 0.075       & 0.037 & 0.042       & **    \\
            & cos(ta)                                         & 0.105       & 0.031 & 0.001       & ***   \\
            & log(sl)                                         & 0.146       & 0.051 & 0.004       & ***   \\
            & cos(ta) : sl                                    & 0.049       & 0.026 & 0.064       & *     \\
            & cos(ta) : log(sl)                               & 0.076       & 0.026 & 0.003       & ***   \\
            & sl : LowActivity                                & -0.917      & 0.113 & \(<\) 0.001 & ***   \\
     \hdashline
     \multirow{5}{*}{Interactions}
            & sl : Water                                      & -0.305      & 0.076 & \(<\) 0.001 & ***   \\
            & sl : Woodland                                   & -0.089      & 0.039 & 0.023       & **    \\
            & sl : Shrubs/Grassland                           & 0.124       & 0.058 & 0.032       & **    \\
            & sl : DistanceToWater \textsuperscript{0.5}      & -0.058      & 0.031 & 0.056       & *     \\
            & cos(ta) : HumanInfluence                        & -0.040      & 0.022 & 0.070       & *     \\
            & cos(ta) : DistanceToWater \textsuperscript{0.5} & 0.063       & 0.026 & 0.017       & **    \\
     \bottomrule
    \end{tabular}
    \begin{tablenotes}
     \item \textit{Significance codes: * \(p < 0.10\) \quad ** \(p < 0.05\)
      \quad *** \(p < 0.01\)}
    \end{tablenotes}
   \end{threeparttable}
  }
 \end{center}
\end{table}

%------------------------------------------------------------------------------
%	Appendix S7: Movement Model interpretation
%------------------------------------------------------------------------------
\newpage
% \newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
\section{Movement Model Interpretation}
To ease with the interpretation of the most parsimonious movement model, we
followed recommendations published in \cite{Fieberg.2021} and produced a series
of plots highlighting how the habitat and movement kernel depended on covariate
values (\Cref{Interpretation}). To visualize the movement kernel and its
interactions with other covariates, we used model estimates and updated our
tentative distribution parameters for turning angles (von Mises distribution
with concentration \(\kappa = 0\)) and step lengths (gamma distribution with
scale \(\theta\) = 6'308 and shape \(k\) = 0.37) by applying the function
\textsf{update\_vonmises()} from the R-package \textsf{amt} \citep{Amt.2019}.
This allowed us to compute probability densities of turning angles and step
lengths under varying values of the associated covariates, while holding all
other covariates constant (\Cref{Interpretation}, a1-a8). Moreover, we
investigated the habitat kernel by computing relative selection strengths (RSS)
between a set of steps where values of the covariate of interest was varied to a
reference step where the covariate value was fixed to its centered value. To
illustrate model uncertainty, we also generated large-sample confidence
intervals using standard errors associated with each model estimate
(\Cref{Interpretation}, b1-b5).

@@

%------------------------------------------------------------------------------
%	Appendix S8: Convergence
%------------------------------------------------------------------------------
\newpage
\section{Convergence}
\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{99_Convergence}
    \caption{Relative traversal frequency through 1'000 checkpoints (5 km x 5
    km) distributed randomly across the study area. The relative traversal
    frequency is plotted against the number of simulated individuals to
    visualize how quickly the metric converges to a steady state. (a) Replicated
    (100 times) relative traversal frequencies across three randomly chosen
    checkpoints as well as the corresponding 95\% prediction interval (PI). (b)
    Averaged relative traversal frequency across all checkpoints and replicates
    including a 95\% PI. (c) Width of the PI in relation to the number of
    simulated dispersers.}
    \label{Convergence}
  \end{center}
\end{figure}

%------------------------------------------------------------------------------
%	Appendix S9: Heatmaps
%------------------------------------------------------------------------------
\newpage
\newgeometry{left=35mm,right=35mm,top=25mm,bottom=5mm,footskip=0pt}
\section{Heatmaps in Relation to the Number of Simulated Steps}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = 0.86\textwidth]{99_HeatmapsIndividual.png}
  \caption{Heatmaps produced when considering 125, 500, and 2000 simulated
   steps, respectively. The left panel (a1, a2, a3) was generated based on
   simulations initiated within the main study area, the right panel (b1, b2,
   b3) was generated based on simulations initiated within the buffer area. To
   produce the heatmap presented in the main manuscript (Figure 5), we tallied
   the values from maps a3 and b3.}
  \label{Heatman}
 \end{center}
\end{figure}
\restoregeometry

%------------------------------------------------------------------------------
%	Appendix S10: Evolution of Betweenness Maps
%------------------------------------------------------------------------------
\newpage
\newgeometry{left=35mm,right=35mm,top=25mm,bottom=10mm,footskip=0pt}
\section{Betweenness Maps in Relation to the Number of Simulated Steps}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = 0.86\textwidth]{99_BetweennessIndividual.png}
  \caption{Maps of betweenness scores produced when considering (a) 125, (b)
   500, (c) and 2000 simulated steps, respectively. A high betweenness score
   indicates that the respective area has a high importance for linking other
   regions in the study area.}
  \label{Betweenness}
 \end{center}
\end{figure}
\restoregeometry

%------------------------------------------------------------------------------
%	Appendix S12: Map Comparison
%------------------------------------------------------------------------------
\newpage
\section{Comparison of Traversal Frequencies and Betweenness Scores Inside and Outside KAZA-TFCA}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = \textwidth]{99_MapComparison.png}
  \caption{Comparison of values from the heatmap and betweenness map inside
  (blue) and outside (orange) the KAZA-TFCA borders.}
  \label{MapComparison}
 \end{center}
\end{figure}

@@

%------------------------------------------------------------------------------
%	Appendix S13: Areas Reached
%------------------------------------------------------------------------------
\newpage
\section{Dispersal into other National Parks}
\begin{figure}[hbtp]
 \begin{center}
  \includegraphics[width = \textwidth]{99_AreasReached.png}
  \caption{Relative number of simulated dispersal trajectories that successfully
  moved from one national park into another that is at least as far away as
  indicated on the x-axis. Percentages are given in relation to the number of
  simulated individuals from the national parks in the respective countries. For
  example, over 85\% of all individuals originating from a national park in
  Angola moved from their natal national park into another one. However, the
  percentage gradually decreases as only national parks at higher euclidean
  distances are considered.}
  \label{AreasReached}
 \end{center}
\end{figure}
